### 2022-08-23 draft/draft.tex
still writing

\newpage
\tableofcontents
\addcontentsline{toc}{section}{\contentsname}

@@

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\setlength{\parindent}{10.0pt}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\argmin}{\mathrm{argmin}}

@@

To estimate $\omega = \begin{bmatrix} \omega_1 & \cdots & \omega_n \end{bmatrix}^\top$ after observing $A$ and supposing that the relative locations of the planets as well as the normalizing constant $C$ are known, one approach might be maximum likelihood maximization. 
For simplicity, set $C = 1$. 
Then the log-likelihood function can be written as:
$$
\ell(\omega) = 
\sum_{i < j} A_{ij} \log \bigg( \frac{\omega_i \omega_j}{d_{ij}^2 -
\omega_i \omega_j} \bigg) + \log (d_{ij}^2 - \omega_i \omega_j) + const.
$$
The partial derivatives with respect to each $\omega_i$ are:
$$
\frac{\partial \ell}{\partial \omega_i} = \sum_{j \neq i} \frac{A_{ij}}{\omega_i} - \frac{\omega_j (1 - A_{ij})}{d_{ij}^2 - \omega_i \omega_j}.
$$
Setting the gradient to zero yields $n$ equations with $n$ unknowns, but it is not entirely obvious how to solve this system of equations. 
Alternatively, we can use gradient ascent or some other numerical optimization method, since this particular log-likelihood function is concave. 
\end{example}

\hypertarget{contributions-of-this-work}{%
\subsection{Contributions of this work}\label{contributions-of-this-work}}

@@

Network analysis is often concerned with community detection.
This has motivated statisticians to develop random graph models with inherent community structure in which the probability of an edge between vertices \(i\) and \(j\) depend on the communities to which vertices \(v_i\) and \(v_j\) belong.
More formally, one assigns each vertex \(v_i\) a community label \(z_i\) and assumes a Bernoulli graph in which \(P_{ij} = g(z_i, z_j, \theta_i, \theta_j)\) for some function \(g: \{1, ..., K\}^2 \times \mathbb{R}^d \times \mathbb{R}^d \to [0, 1]\) and \(\theta_i \in \mathbb{R}^d\) is a vector of real-valued parameters associated with each vertex.
Within the context of this work, we will call such models \emph{block models}.
This type of model frmaework allows us to frame the goal of community detection as a statistical inference problem:
identify the true community labels, up to permutation of the labels.

@@

The SBM is a type of mixture model, and like most classical mixture models and clustering problems, likelihood maximization is NP-hard.
Because the SBM is a mixture model, a candidate algorithm for finding a (local) maximum of the likelihood is the expectation maximization (EM) algorithm \citep{10.2307/2984875}.
However, due to the \(z_{ik} z_{j \ell}\) interaction terms, the expectation step cannot be solved in closed form \citep{kolaczyk2014statistical}.
While the labels are drawn independently, they are not necessarily conditionally independent.
Nevertheless, making the conditional independence relaxation allows us to write the expectation and maximization steps in closed form.
As an example, we write the EM algorithm for the likelihood function in equation \ref{eq:sbm-loglik} in algorithm \ref{algo:sbm-em}.

\begin{algorithm}
\label{algo:sbm-em}
  \DontPrintSemicolon
  \SetAlgoLined
  \KwData{Adjacency matrix $A$, number of communities $K$}
  \KwResult{Estimated community label probabilities $\{\pi_{ik}\}$, estimated community edge probabilities $\{\hat{\theta}_{k \ell}\}$}
  \caption{Approximate EM algorithm for the SBM}
  Initialize $\{\pi_{ik}\}$, $\{\theta_{k \ell}\}$.\;
  \While{$\|\nabla \ell\| > \epsilon$} {
    \For {$i = 1, ..., n$} {
      \For {$k = 1, ..., K$} {
        E-step: $\pi_{ik} \propto \exp \bigg( \sum_{j \neq i} \sum_{\ell} \pi_{j \ell} (A_{ij} \log \theta_{k \ell} + (1 - A_{ij}) \log (1 - \theta_{k \ell})) \bigg)$.\;
        M-step: $\theta_{k \ell} = \frac{\sum_{i < j} A_{ij} \pi_{ik} \pi_{j \ell}}{\sum_{i < j} \pi_{ik} \pi_{j \ell}}$.\;
      }
    }
  }
\end{algorithm}

@@

\hypertarget{random-graph-models-and-sparsity}{%
\subsubsection{Random Graph Models and Sparsity}\label{random-graph-models-and-sparsity}}

@@

\hypertarget{the-hierarchy-of-block-models}{%
\subsubsection{The Hierarchy of Block Models}\label{the-hierarchy-of-block-models}}

@@

\hypertarget{sec:sbm-dcbm-grdpg}{%
\subsection{Connecting the Stochastic Block Model and the Degree Corrected Block Model to the Generalized Random Dot Product Graph}\label{sec:sbm-dcbm-grdpg}}

@@

However, because we do not observe \(P\) but rather only the noisy
adjacency matrix \(A \sim \BG(P)\), the natural approach then is
to perform SSC on the rows of the spectral embedding of \(A\),
since the embedding of \(P\) consists of \(K\) subspaces (theore \ref{thm:pabm-grdpg}),
and so the embedding of \(A\) will lie approximately on the \(K\) subspaces.
We will show in theorem \ref{theorem5} that, with probability converging to
one as \(n \rightarrow \infty\), the rows of the ASE of \(A\) also
satisfy the subspace detection property.
Theorem \ref{theorem5} builds upon existing work by
\citet{rubindelanchy2017statistical} who describe the convergence
behavior of the ASE of \(A\) to that of \(\tilde{\Pi} XU\), and
\citet{jmlr-v28-wang13} who show the necessary conditions for the
subspace detection property to hold in noisy cases where the points lie
near subspaces. Finally we emphasize that while
\citet{noroozi2019estimation} also considered the use of SSC for
community recovery in PABM, they instead applied SSC to the rows of
\(A\) itself, foregoing
the embedding step altogether. It is however much harder to show that
the rows of \(A\) satisfy the subspace detection property and thus, to
the best of our knowledge, there is currently no consistency result
regarding the application of SSC to the rows of \(A\).

@@

@@

\hypertarget{generalized-random-dot-product-graphs-with-community-structure}{%
\section{Generalized Random Dot Product Graphs with Community Structure}\label{generalized-random-dot-product-graphs-with-community-structure}}

@@

>>>>

% \tableofcontents

\addcontentsline{toc}{section}{\contentsname}

@@

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{2}
\tableofcontents
}
\setlength{\parindent}{20.0pt}
\newcommand{\diag}{\mathrm{diag}}
\newcommand{\tr}{\mathrm{Tr}}
\newcommand{\blockdiag}{\mathrm{blockdiag}}
\newcommand{\indep}{\stackrel{\mathrm{ind}}{\sim}}
\newcommand{\iid}{\stackrel{\mathrm{iid}}{\sim}}
\newcommand{\Bernoulli}{\mathrm{Bernoulli}}
\newcommand{\Betadist}{\mathrm{Beta}}
\newcommand{\BG}{\mathrm{BernoulliGraph}}
\newcommand{\Uniform}{\mathrm{Uniform}}
\newcommand{\PABM}{\mathrm{PABM}}
\newcommand{\RDPG}{\mathrm{RDPG}}
\newcommand{\GRDPG}{\mathrm{GRDPG}}
\newcommand{\Multinomial}{\mathrm{Multinomial}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\as}{\stackrel{\mathrm{a.s.}}{\to}}
\newcommand{\ER}{\text{Erd\"{o}s-R\'{e}nyi}}
\newcommand{\SBM}{\mathrm{SBM}}
\newcommand{\DCBM}{\mathrm{DCBM}}
\newcommand{\argmin}{\mathrm{argmin}}

@@

To estimate $\omega = \begin{bmatrix} \omega_1 & \cdots & \omega_n \end{bmatrix}^\top$ after observing $A$ and supposing that the relative locations of the planets as well as the normalizing constant $C$ are known, one approach might be maximum likelihood maximization. 
For simplicity, set $C = 1$. 
Then the log-likelihood function can be written as:
$$
\ell(\omega; A) = 
\sum_{i < j} A_{ij} \log \bigg( \frac{\omega_i \omega_j}{d_{ij}^2 -
\omega_i \omega_j} \bigg) + \log (d_{ij}^2 - \omega_i \omega_j) + const.
$$
The partial derivatives with respect to each $\omega_i$ are:
$$
\frac{\partial \ell}{\partial \omega_i} = \sum_{j \neq i} \frac{A_{ij}}{\omega_i} - \frac{\omega_j (1 - A_{ij})}{d_{ij}^2 - \omega_i \omega_j}.
$$
Setting the gradient to zero yields $n$ equations with $n$ unknowns, but it is not entirely obvious how to solve this system of equations. 
Alternatively, we can use gradient ascent or some other numerical optimization method, since this particular log-likelihood function is concave. 
\end{example}

\hypertarget{contributions-of-this-work}{%
\subsection{Contributions of This Work}\label{contributions-of-this-work}}

@@

Network analysis is often concerned with community detection.
This has motivated statisticians to develop random graph models with inherent community structure in which the probability of an edge between vertices \(i\) and \(j\) depend on the communities to which vertices \(v_i\) and \(v_j\) belong.
More formally, one assigns each vertex \(v_i\) a community label \(z_i\) and assumes a Bernoulli graph in which \(P_{ij} = g(z_i, z_j, \theta_i, \theta_j)\) for some function \(g: \{1, ..., K\}^2 \times \mathbb{R}^d \times \mathbb{R}^d \to [0, 1]\) and \(\theta_i \in \mathbb{R}^d\) is a vector of real-valued parameters associated with each vertex.
Within the context of this work, we will call such models \emph{block models}.
This type of model framework allows us to frame the goal of community detection as a statistical inference problem:
identify the true community labels, up to permutation of the labels.

@@

The SBM is a type of mixture model, and like most classical mixture models and clustering problems, likelihood maximization is NP-hard.
Because the SBM is a mixture model, a candidate algorithm for finding a (local) maximum of the likelihood is the expectation maximization (EM) algorithm \citep{10.2307/2984875}.
However, due to the \(z_{ik} z_{j \ell}\) interaction terms, the expectation step cannot be solved in closed form \citep{kolaczyk2014statistical}.
While the labels are drawn independently, they are not necessarily conditionally independent.
Nevertheless, making the conditional independence relaxation allows us to write the expectation and maximization steps in closed form.
As an example, we write the EM algorithm for the likelihood function in equation \eqref{eq:sbm-loglik} in algorithm \ref{algo:sbm-em}.

\begin{algorithm}
\label{algo:sbm-em}
  \setstretch{1.5}
  \DontPrintSemicolon
  \SetAlgoLined
  \KwData{Adjacency matrix $A$, number of communities $K$}
  \KwResult{Estimated community label probabilities $\{\pi_{ik}\}$, estimated community edge probabilities $\{\hat{\theta}_{k \ell}\}$}
  \caption{Approximate EM algorithm for the SBM}
  Initialize $\{\pi_{ik}\}$, $\{\theta_{k \ell}\}$.\;
  \While{$\|\nabla \ell\| > \epsilon$} {
    \For {$i = 1, ..., n$} {
      \For {$k = 1, ..., K$} {
        E-step: $\pi_{ik} \propto \exp \bigg( \sum_{j \neq i} \sum_{\ell} \pi_{j \ell} (A_{ij} \log \theta_{k \ell} + (1 - A_{ij}) \log (1 - \theta_{k \ell})) \bigg)$.\;
        M-step: $\theta_{k \ell} = \frac{\sum_{i < j} A_{ij} \pi_{ik} \pi_{j \ell}}{\sum_{i < j} \pi_{ik} \pi_{j \ell}}$.\;
      }
    }
  }
\end{algorithm}

@@

\hypertarget{random-graph-models-and-sparsity}{%
\subsection{Random Graph Models and Sparsity}\label{random-graph-models-and-sparsity}}

@@

\hypertarget{the-hierarchy-of-block-models}{%
\subsection{The Hierarchy of Block Models}\label{the-hierarchy-of-block-models}}

@@

\hypertarget{sec:sbm-dcbm-grdpg}{%
\subsection{Connecting the SBM and DCBM to the GRDPG}\label{sec:sbm-dcbm-grdpg}}

@@

However, because we do not observe \(P\) but rather only the noisy
adjacency matrix \(A \sim \BG(P)\), the natural approach then is
to perform SSC on the rows of the spectral embedding of \(A\),
since the embedding of \(P\) consists of \(K\) subspaces (theore \ref{thm:pabm-grdpg}),
and so the embedding of \(A\) will lie approximately on the \(K\) subspaces.
We will show in theorem \ref{thm:ssc} that, with probability converging to
one as \(n \rightarrow \infty\), the rows of the ASE of \(A\) also
satisfy the subspace detection property.
Theorem \ref{thm:ssc} builds upon existing work by
\citet{rubindelanchy2017statistical} who describe the convergence
behavior of the ASE of \(A\) to that of \(\tilde{\Pi} XU\), and
\citet{jmlr-v28-wang13} who show the necessary conditions for the
subspace detection property to hold in noisy cases where the points lie
near subspaces. Finally we emphasize that while
\citet{noroozi2019estimation} also considered the use of SSC for
community recovery in PABM, they instead applied SSC to the rows of
\(A\) itself, foregoing
the embedding step altogether. It is however much harder to show that
the rows of \(A\) satisfy the subspace detection property and thus, to
the best of our knowledge, there is currently no consistency result
regarding the application of SSC to the rows of \(A\).

\hypertarget{algorithms-for-community-detection}{%
\subsubsection{Algorithms for Community Detection}\label{algorithms-for-community-detection}}

@@

Theorem \ref{thm:osc} guarantees that for any \(\epsilon > 0\),
the number of edges of \(\hat{B}\) between vertices of different communities that are larger than \(\epsilon\) converges to zero with probability converging to one as \(n\) increases.
We can always find an \(\epsilon > 0\) such that \(\hat{B}_{ij} > \epsilon\) with probability converging to one as \(n\) increases.
Thus, by using \(\hat{B}\), we can perfectly recover all the latent community assignments \(z_1, z_2, \dots, z_n\), i.e., the number of misclustered vertices is zero asymptotically almost surely.
We note that theorem \ref{thm:osc} is stronger than existing results in the literature;
in particular theorem 1 of \citet{307cbeb9b1be48299388437423d94bf1} (the paper that
originally introduces the PABM model) only guarantees that the \emph{proportion} of misclustered vertices converges to \(0\) as \(n \rightarrow \infty\).
Furthermore theorem 1 of \citet{307cbeb9b1be48299388437423d94bf1} also requires the sparsity parameter \(\rho_n\) to satisfies \(n \rho_n^2 = \omega(\log^2{n})\) which is a considerably stronger assumption than the assumption \(n \rho_n =\omega(\log^{4}{n})\) used in theorem \ref{thm:osc}.
Indeed, \(n\rho_n^2 = \omega(\log^{2}{n})\) implies \(n \rho_n = \omega(n^{1/2})\).
We emphasize that the assumption \(n \rho_n = \omega(\log^{c}{n})\) for some constant \(c > 1\) is commonly used in the context of graph estimation using spectral methods.

Theorems \ref{thm:pabm-grdpg}, \ref{thm:osc-p}, and \ref{thm:osc} also provide
a natural path toward using SSC for community detection.
In particular we established in theorem \ref{thm:pabm-grdpg} that an ASE of the edge
probability matrix \(P\) can be constructed from a latent vector configuration
consisting of orthogonal subspaces. Theorem \ref{thm:osc-p} shows how
this property can also be recovered from the eigenvectors of \(P\).
Then theorem \ref{thm:osc} shows that, by replacing \(P\) with \(A\), the
rows of \(\hat{V}\) also lie on asymptotically orthogonal subspaces.
Motivated by theorem \ref{thm:osc}, theorem \ref{thm:ssc} below
shows that the subspace detection property also holds for the rows of
\(\sqrt{n} \hat{V}\).

\begin{algorithm}[t]
  \label{alg:ssc}
  \setstretch{1.5}
  \DontPrintSemicolon
  \SetAlgoLined
  \caption{Sparse Subspace Clustering using LASSO.}
  \KwData{Adjacency matrix $A$, number of communities $K$,
  hyperparameter $\lambda$}
  \KwResult{Community assignments $z_1, ..., z_n \in \{1, ..., K\}$}
    Find $V$, the matrix of eigenvectors of $A$
    corresponding to the $K (K + 1) / 2$ most positive
    and the $K (K - 1) / 2$ most negative eigenvalues.\;
    Normalize $V \leftarrow \sqrt{n} V$.\;
    \For {$i = 1, ..., n$} {
      Assign $v_i^\top$ as the $i^{th}$ row of $V$.
      Assign $V_{-i} = \bigl[
      v_1 \mid \cdots \mid v_{i-1} \mid v_{i+1} \mid \cdots \mid v_n \bigr]^\top$.\;
      Solve the LASSO problem
      $c_i = \arg\min_{\beta}
      \frac{1}{2} \|v_i - V_{-i} \beta\|_2^2 + \lambda \|\beta\|_1$.\;
      Assign $\tilde{c}_i = (c_i^{(1)}, \dots, c_i^{(i-1)}, 0, c_i^{(i)}, \dots, c_i^{(n-1)})^\top$ such that the superscript is the index of
      $\tilde{c}_i$.\;
    }
    Assign
    $C = \bigr[ \tilde{c}_1 \mid \cdots \mid \tilde{c}_n \bigr]$.\;
    Compute the affinity matrix $B = |C| + |C^\top|$.\;
    Construct graph $G$ using $B$ as its similarity matrix.\;
    Partition $G$ into $K$ disconnected subgraphs (e.g., using edge
    thresholding or spectral clustering).\;
    Map each partition to the community labels $1, ..., K$.
\end{algorithm}

\begin{theorem}
\label{thm:ssc}
Let $P$ describe the edge probability matrix of the PABM with $n$ vertices, and let $A \sim \Bernoulli(P)$. 
Let $\hat{V}$ be the matrix of eigenvectors of $A$ corresponding to the $K^2$ largest eigenvalues in modulus. 
Then for any $\epsilon > 0$, there exists a choice of $\vartheta > 0$ and $N \in \mathbb{N}$ such that for all $n \geq N$, $\sqrt{n} \hat{V}$ obeys the subspace detection property with probability at least $1 - \epsilon$.  
\end{theorem}

\hypertarget{algorithm-for-parameter-estimation}{%
\subsubsection{Algorithm for Parameter Estimation}\label{algorithm-for-parameter-estimation}}

For ease of exposition we now assume in this subsection that the edge probability matrix \(P\) for the PABM had been arranged so that the rows and columns are organized by community so that \(\tilde{P} = P\) (see remark \ref{rem:pabm_view2}).
Then the \(k\ell\)\textsuperscript{th} block is an outer product of two vectors, i.e., \(P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^\top\).
Therefore, given \(P^{(k \ell)}\), \(\lambda^{(k \ell)}\) and \(\lambda^{(\ell k)}\) are solvable up to multiplicative constant using singular value decomposition.
More specifically, let \(P^{(k \ell)} = (\sigma^{(k \ell)})^2 u^{(k \ell)} (v^{(k \ell)})^\top\) be the singular value decomposition of \(P^{(k \ell)}\) where \(u^{(k \ell)} \in \mathbb{R}^{n_k}\) and \(v^{(k \ell)} \in \mathbb{R}^{n_\ell}\) are vectors and \(\sigma^{(k \ell)}\) is a scalar.
Then \(\rho_n^{1/2} \lambda^{(k \ell)} = s_1 u^{(k \ell)}\) and \(\rho_n^{1/2} \lambda^{(\ell k)} = s_2 v^{(k \ell)}\) for unidentifiable \(s_1 s_2 = (\sigma^{(k \ell)})^2\).
Because each \(\lambda^{(k \ell)}\) is not strictly identifiable, we instead estimate each \(\tilde{\lambda}^{(k \ell)} = \sigma^{(k \ell)} u^{(k \ell)}\).
Given the adjacency matrix \(A\) instead of edge probability matrix \(P\), we can simply use plug-in estimators by taking the SVD of each \(A^{(k \ell)}\) to obtain \(\hat{\lambda}^{(k \ell)} = \hat{\sigma}^{(k \ell)} \hat{u}^{(k \ell)}\) using the largest singular value of \(A\) and its corresponding singular vectors.

\begin{algorithm}[tp]
  \label{alg:param_est}
  \DontPrintSemicolon
  \SetAlgoLined
  \setstretch{1.5}
  \caption{PABM parameter estimation.}
  \KwData{Adjacency matrix $A$, community assignments $1, ..., K$}
  \KwResult{PABM parameter estimates $\{\hat{\lambda}^{(k \ell)}\}_K$.}
  Arrange the rows and columns of $A$ by community such that each 
  $A^{(k \ell)}$ block consists of estimated edge probabilities between 
  communities $k$ and $l$.\;
  \For {$k, \ell = 1, ..., K$, $k \leq \ell$} {
    Compute $A^{(k \ell)} = U \Sigma V^\top$, the SVD of the $k\ell^{th}$ block.\;
    Assign $u^{(k \ell)}$ and $v^{(k \ell)}$ as the first columns of $U$ and $V$. 
    Assign $(\sigma^{(k \ell)})^2 \leftarrow \Sigma_{11}$.\;
    Assign $\hat{\lambda}^{(k \ell)} \leftarrow \pm \sigma^{(k \ell)} u^{(k \ell)}$ and $\hat{\lambda}^{(\ell k)} \leftarrow \pm \sigma^{(k \ell)} v^{(k \ell)}$.
  }
\end{algorithm}

\begin{theorem}
\label{thm:param-est}
Let each $\tilde{\lambda}^{(k \ell)}$ be the popularity vector derived from its corresponding $P^{(k \ell)}$ and let $\hat{\lambda}^{(k \ell)}$ be its estimate obtained from $A^{(k \ell)}$ using algorithm~\ref{alg:param_est}.
Then if $n \rho_n = \omega( \log^{4}{n})$,
\begin{equation} 
\label{eq:param-est}
\max_{k, \ell \in \{1, ..., K\}} 
\|\hat{\lambda}^{(k \ell)} - \tilde{\lambda}^{(k \ell)}\|_{\infty} = 
O\bigg(\frac{\log n_k}{\sqrt{n_k}} \bigg)
\end{equation}
with high probability. 
Here $\|\cdot\|_{\infty}$ denotes the $\ell_\infty$ norm of a vector. Let $\hat{\Lambda}$ be the matrix
$$\hat{\Lambda} = \begin{bmatrix} \hat{\lambda}^{(11)} &
  \hat{\lambda}^{(12)} & \cdots & \hat{\lambda}^{(1 K)} \\
  \hat{\lambda}^{(21)} &
  \hat{\lambda}^{(22)} & \cdots & \hat{\lambda}^{(2 K)} \\
  \vdots & \vdots & \cdots & \vdots \\
   \hat{\lambda}^{(K1)} &
  \hat{\lambda}^{(K2)} & \cdots & \hat{\lambda}^{(K K)} 
\end{bmatrix}$$
and let $\hat{P} = \hat{X} U I_{p,q} U^\top \hat{X}^{\top}$, where $\hat{X}$ is defined from $\hat{\Lambda}$ and $U$ is defined from $K$ as in theorem~\ref{thm:pabm-grdpg}. 
Equation~\eqref{eq:param-est} then implies
\begin{equation}
\label{eq:conv_prob}
\frac{1}{n} \|\rho_n^{-1} \hat{P} - \rho_n^{-1} P\|_{F} = O((n
\rho_n)^{-1/2}), \quad \max_{ij} |\rho_n^{-1} \hat{P}_{ij} -
\rho_n^{-1} P_{ij}| = O((n \rho_n)^{-1/2})
\end{equation}
with high probability. 
\end{theorem}

Equation \eqref{eq:param-est} guarantees that \(n^{-1/2} \|\rho_n^{-1/2} \hat{\Lambda} - \Lambda\|_{F} = O((n \rho_n)^{-1/2})\).
Equation \eqref{eq:conv_prob} then guarantees that the mean square error for \(\rho_n^{-1} (\hat{P} - P)\) converges to \(0\) almost surely and furthermore the entries of \(\rho_n^{-1} \hat{P}\) converge uniformly to the entries of \(\rho_n^{-1} P\);
recall that \(\rho_n^{-1} P_{ij} = \lambda_{iz_j} \lambda_{j z_i}\).
We note that these results are stronger than existing results in \citet{307cbeb9b1be48299388437423d94bf1};
for example theorem 2 in \citet{307cbeb9b1be48299388437423d94bf1} only guarantees
\(n^{-1/2} \|\rho_n^{-1/2} \hat{\Lambda} - \Lambda\|_{F} = o(1)\) as \(n \rightarrow \infty\).

\hypertarget{simulation-study}{%
\subsection{Simulation Study}\label{simulation-study}}

\hypertarget{applications}{%
\subsection{Applications}\label{applications}}

@@

\hypertarget{generalized-random-dot-product-graphs-with-nonlinear-community-structure}{%
\section{Generalized Random Dot Product Graphs with Nonlinear Community Structure}\label{generalized-random-dot-product-graphs-with-nonlinear-community-structure}}

\hypertarget{the-manifold-block-model}{%
\subsection{The Manifold Block Model}\label{the-manifold-block-model}}

\hypertarget{algorithms-for-nonintersecting-manifolds}{%
\subsection{Algorithms for Nonintersecting Manifolds}\label{algorithms-for-nonintersecting-manifolds}}

\hypertarget{algorithms-for-intersecting-manifolds}{%
\subsection{Algorithms for Intersecting Manifolds}\label{algorithms-for-intersecting-manifolds}}

\hypertarget{examples-and-simulations}{%
\subsection{Examples and Simulations}\label{examples-and-simulations}}

\hypertarget{applications-1}{%
\subsection{Applications}\label{applications-1}}

\hypertarget{conclusions}{%
\subsection{Conclusions}\label{conclusions}}

@@

We now provide a proof of theorem \ref{thm:ssc}. Our proof is based on verifying the sufficient conditions given in theorem 6 of \citet{jmlr-v28-wang13} under which sparse subspace clustering based on solving the optimization problem in equation \eqref{eq:ssc} yields an affinity matrix \(B = |C| + |C^{\top}|\) satisfying the subspace detection property of definition \ref{def:subspace_detection}.
We first recall a few definitions used in \citet{soltanolkotabi2012} and \citet{jmlr-v28-wang13};
for ease of exposition, these definitions are stated using the notations of the current paper and we will drop the explicit dependency on \(n\) from our eigenvectors \(\hat{V}\) of \(A\) and \(V\) of \(P\).

\begin{definition}[Inradius]
\label{def:inradius}
The inradius of a convex body $\mathcal{P}$, denoted by $r(\mathcal{P})$, is
defined as the radius of the largest Euclidean ball inscribed in $\mathcal{P}$.
Let $X$ be a $n \times d$ matrix with rows $x_1, x_2, \dots,
x_n$. We then define, with a slight abuse of notation, $r(X)$ as the
inradius of the convex hull formed by $\{\pm x_1, \pm x_2, \dots, \pm x_n\}$. 
\end{definition}

\begin{definition}[Subspace incoherence]
\label{def:subspace_incoherence}
Let $\hat{V}$ be the eigenvectors of $A$
corresponding to the $K^2$ largest eigenvalues in modulus. Let
$\hat{V}^{(k)}$ denote the matrix formed by keeping only the rows of
$\hat{V}$ corresponding to the $k^{th}$
community and let $\hat{V}^{(-k)}$ denote the matrix formed by
omitting the rows of $\hat{V}$
corresponding to the $k^{th}$ community. Let $(\hat{v}_i^{(k)})^\top$ denote
the $i$th row of $\hat{V}^{(k)}$ and $\hat{V}_{-i}^{(k)}$ be $\hat{V}^{(k)}$ with
the $i^{th}$ row omitted. Let $V$, $V^{(k)}$, $V^{(-k)}$, and
$v_i^{(k)}$ be defined similarly using the eigenvectors $V$ of
$P$. Finally let $\mathcal{S}^{(k)}$ be the vector space spanned by the
rows of $V^{(k)}$.

Now define $\nu_{i}^{(k)}$ for $k = 1,2,\dots,K$ and $i =
1,2,\dots,n_{k}$ as the solution of the following optimization problem
$$\nu_{i}^{(k)} = \max_\eta (\hat{v}_i^{(k)})^\top \eta - \frac{1}{2
  \lambda} \eta^\top \eta, \quad \text{subject to $\|V_{-i}^{(k)}
  \eta\|_\infty \leq 1$.}$$
Given $\nu_i^{(k)}$, let $\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})$
be the vector in $\mathbb{R}^{K^2}$ corresponding to the orthogonal projection of $\nu_i^{(k)}$ onto the vector space
$\mathcal{S}^{(k)}$ and define the projected dual direction $w_{i}^{(k)}$
as
$$w_i^{(k)} =
\frac{\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})}{\|\mathbb{P}_{\mathcal{S}^{(k)}}(\nu_i^{(k)})\|}.$$
Now let $W^{(k)} = \bigl[ w_1^{(k)} \mid \cdots \mid w_{n_k}^{(k)} \bigr]^\top$
and define the subspace incoherence for $\hat{V}^{(k)}$ by
$$\mu^{(k)} = \mu(\hat{V}^{(k)}) = \max\limits_{v \in V^{(-k)}} \|W^{(k)} v\|_\infty.$$
\end{definition}

\begin{proof}[Proof of theorem \ref{thm:ssc}]
For a given $k = 1,2\dots,K$, let $r^{(k)} = \min_{i}r(V_{-i}^{(k)})$ be inradius of the convex hull formed by
the rows of $V_{-i}^{(k)}$ and let $r_* = \min_{k} r^{(k)}$. Then Theorem~6 in
\citet{jmlr-v28-wang13} states that there exists a $\lambda > 0$
such that $\sqrt{n} \hat{V}$ satisfies the subspace detection property in definition \ref{def:subspace_detection} whenever the following two conditions are satisfied: 
\begin{gather}
  \label{eq:cond1}
  \mu^{(k)} < r^{(k)} \quad \text{for all $k = 1,2,\dots,K$}, \\
  \label{eq:cond2}
  \max_{i} \|W \hat{v}_{i} - v_{i}\| \leq \min_{k} \frac{r_*(r^{(k)} -
    \mu^{(k)})}{2 + 7 r^{(k)}}.
\end{gather}
We now verify that for sufficiently large n, equation \eqref{eq:cond1} and Eq.~\eqref{eq:cond2}
holds with high probability.

{\bf Verifying equation \eqref{eq:cond1}}. If $n$ is sufficiently large then
there are enough vertices in each community $k$ so that
$\mathrm{span}(V_{-i}^{(k)}) = \mathcal{S}^{(k)}$ for all $i$ and hence
\(r^{(k)} = \min_{i} r(V_{-i}^{(k)}) > 0\) for
all $k = 1,2,\dots,K$.

Next, by theorem \ref{thm:osc-p} we have that the subspaces
$\{\mathcal{S}^{(1)}, \dots, \mathcal{S}^{(K)}\}$
are mutually orthogonal, i.e., $v^{\top} w = 0$ for all $v \in
\mathcal{S}^{(k)}$ and $w \in \mathcal{S}^{(\ell)}$ with $k \not =
\ell$. Now let $z \in \mathbb{R}^{K^2}$ be arbitrary and let
$\tilde{z} = \mathbb{P}_{\mathcal{S}^{(k)}} z$ be the projection of
$z$ onto $\mathcal{S}^{(k)}$. We then have $v^{\top} \tilde{z} =
0$ for all $v \in V^{(-k)}$. Because $z$ is arbitrary, this implies 
$\|W^{(k)} v\|_{\infty} = 0$ for all $v
\in V^{(-k)}$ and hence $\mu^{(k)} = 0$ for all $k
=1,2,\dots,K$. Therefore $\mu^{(k)} < r^{(k)}$ for all $k =
1=2,\dots,K$ as desired.

{\bf Verifying equation \eqref{eq:cond2}}.
Let $\delta = \max_{i} \sqrt{n} \|W \hat{v}_{i} - v_{i}\|$. 
Then from lemma \ref{lem:technical}, we have \(\delta \stackrel{a.s.}{\to} 0\) and hence
$$\delta < \min_{k} \frac{r_* (r^{(k)} - \mu^{(k)})}{2 + 7 r^{(k)}}$$
asymptotically almost surely.

In summary $\sqrt{n} \hat{V}$ satisfies the subspace detection property with probability converging to $1$ as \(n \to \infty\).
\end{proof}

\begin{remark}
Theorem 6 of \citet{jmlr-v28-wang13} assumes that each row $v_i$
of $V$ has unit norm, i.e., $\|v_{i}\| = 1$  for all $i$. 
This assumption has the effect of scaling the $r^{(k)}$ so that $r^{(k)} \leq 1$ for all $k = 1,2,\dots,K$. 
We emphasize that this assumption has no effect on the proof of Theorem \ref{thm:ssc}. 
Indeed, because $\mu^{(k)} = 0$ for all $k$, as long as the rows of $V^{(k)}$ spans the subspace $\mathcal{S}^{(k)}$, then $a r^{(k)} > \mu^{(k)}$ for any scalar $a > 0$. 
\end{remark}

\begin{proof}[Proof of Theorem \ref{thm:param-est}] 
Let \(P\) be organized by community such that \(P^{(k \ell)}\) denote the $n_k \times n_{\ell}$ matrix obtained by keeping only the rows of $P$ corresponding to vertices in community $k$ and the columns of $P$ corresponding to vertices in community $\ell$. 
We define $A^{(k \ell)}$ analogously. 
Recall that $P^{(k \ell)} = \lambda^{(k \ell)} (\lambda^{(\ell k)})^{\top}$ for all $k, \ell$. 
We now consider estimation of $P^{(k \ell)}$ for the cases when $k = \ell$ versus when $k \not = \ell$.
